### 2019-09-21 190904_Paper_Thesis/Conclusion.tex
Paper_Thesis

\label{chp:Conclusions}
\section{Conclusions}
...

\section{Future Works}
...
\begin{enumerate}
	\setlength{\itemsep}{1pt}
	\item Recognizing the in-air signature written along non-LOS. 
	\item Repeating the same data acquisition experiments in a different place
	\item Studying identification performances when multiple people are staying in the same room.
	\item To secure a reliable identification system, spoofing attack scenario should be investigated. 
\end{enumerate}