### 2019-09-21 190904_Paper_Thesis/Experiments.tex

\label{chp:Experiments}
..
\section{Data set}
\label{sec:Dataset}
..
\section{Experimental Settings}
..
\subsection{Evaluation Scenarios}
\label{sec:scenario}
..
\subsection{Evaluation Protocols}
\label{sec:protocols}
..
\subsection{Parameter Settings}
\label{sec:parameter}
..
\section{Experimental Results}
\label{sec:results}
..
\subsection{Effect of the DC Component}
\label{sec:DC}
..
\subsection{Effect of the Weighting Scheme}
..
\subsection{Effect of the Facing Direction and the Position of the Subject}
..
\subsection{Comparison with the ConvNet}
..