### 2019-09-21 190904_Paper_Thesis/Preliminaries.tex

\label{chp:Preliminary}
..
\section{Biometric Identification System}
\label{sec:Biometrics}
..
\section{Evaluation metrics}
\label{sec:Evaluation}
..
\section{Received Signal Strength Indicator}
\label{sec:RSSI}
..
\section{Channel State Information}
\label{sec:CSI}
..
\subsection{Multiple-Input and Multiple-Output}
\label{sec:MIMO}
..
\section{Shift Matrix}
\label{sec:Shift}
..
\section{Metric Learning}
\label{sec:metric_learning}