### 2019-09-21 190904_Paper_Thesis/abstract.tex

\begin{center}
	\LARGE An In-Air Signature Identification System using Commercial Wi-Fi Devices
\end{center}

\parbox[t]{0.4\textwidth}

{
	\begin{flushright}
		Han-Cheol Moon\\
		School of Electrical and Electronic Engineering\\
		The Graduate School\\
		Yonsei University
	\end{flushright}
}
\vspace{0.1em}

...

\vspace{\stretch{1}}
\noindent
\hrulefill\\
{\bf Key words : Biometrics, Identification, In-Air Signature, Channel State Information, and
	Wi-Fi}
\parbox[t]{0.8\textwidth}
{}