### 2019-09-27 190904_Paper_Thesis/Introduction.tex

\section{Background}
% biometric
As mobile devices increase, identity verification using biometrics traits is becoming more widely used.
Compared to the use of complex passwords, using biometric systems to user verification has advantages of increased security, convinience and accountability~\ref{hutton2004biometrics}.

among the biometric traits, signatures have also been widely used as a way of identification [].
With the development of various sensors such as depth camera and mobile camera, in-air signature [] have also become available.

A system has been developed to recognize signatures written in the air.
- Easy to use and no tools required
Signatures written in Wi-Fi CSI signals are also recognizable
- Low accuracy and difficult recognition of signatures in different directions
- Due to the nature of the hollow signature, it is important to recognize the different directions of signature.
We are trying to solve this problem by using deep learning.

%\newpage
\section{Motivation and Contributions}
...
\newpage
\section{Organization of Thesis}
..