### 2019-10-07 190904_Paper_Thesis/Conclusion.tex
paper_thesis:v0

\label{chp:Conclusions}
\section{Conclusions}
In this paper, we proposed a system for identity verification from Wi-Fi in-air signature signals based on triplet network. Three-channel ConvNet structures was adopted in order to learn discriminative features from the in-air signatures. Moreover, we proposed a input triplet mining approach based on the kernel and range space learning to faster the convergence speed.
Our experimental results on in-house Wi-Fi handwritten signature dataset showed the proposed network outperforms handcraft methods and Siamese network by improved verification accuracy and faster loss convergence.
\iffalse
\section{Future Works}
\begin{enumerate}
	\setlength{\itemsep}{1pt}
	\item Recognizing the in-air signature written along non-LOS. 
	\item Repeating the same data acquisition experiments in a different place
	\item Studying identification performances when multiple people are staying in the same room.
	\item To secure a reliable identification system, spoofing attack scenario should be investigated. 
\end{enumerate} 
\fi