### 2019-10-17 191016_Paper_Thesis_v2/thanks.tex

\newpage
{\LARGE \bf
\noindent ACKNOWLEDGEMENT}\\[11pt]
\vskip 1em \pagestyle{empty}
\renewcommand{\baselinestretch}{0.85}

\begin{flushright}
{October 2019 \\ \textbf{Young-Woong Kwon}}
\renewcommand{\baselinestretch}{1.0}
\end{flushright}
First of all, I would like to thank Professor Kar-Ann Toh for steering the direction of my research. His warm encouragement and careful guidance were very helpful for me.
Thank you for my director Kiheon Kim and my manager Jongman Kim for recommending me to study at graduate school.
Thanks for my company, Samsung SDI for supporting me to study at graduate school.
I also want to give my special thanks to Jooyoung, Kim. His consideration and positive attitude always motivated me to do better.
Thank you to all my lab members for helping to enjoy school life.
Last but not least, I couldn't have finished this work without my dear wife, who always trusted and waited for me despite going back to graduate school at a late age.