### 2019-11-06 191016_Paper_Thesis_v2/Conclusion.tex
paper_thesis:update

\label{chp:Conclusion}
%\section{Conclusion}
In this thesis, a system for identity verification was developed based on the hand gesture signature signals sensed by the Wi-Fi Channel State Information (CSI).
A three-channel ConvNet structure was adopted to learn the discriminative features based on a relatively small size in-air handwritten signature dataset. We proposed an input triplet mining based on the kernel and range space learning to improve the convergence speed of the triplet network training.
Our experimental results on the Wi-Fi CSI signature dataset showed encouraging accuracy and convergence performances.

\iffalse
\section{Future Work}
\begin{enumerate}
	\setlength{\itemsep}{1pt}
	\item Recognizing the in-air signature written along non-LOS. 
	\item Repeating the same data acquisition experiments in a different place
	\item Studying identification performances when multiple people are staying in the same room.
	\item To secure a reliable identification system, spoofing attack scenario should be investigated. 
\end{enumerate} 
\fi