### 2019-11-06 191016_Paper_Thesis_v2/Introduction.tex

\section{Background}
% signature
Handwritten signatures have long been used as an identification tool and various methods for verifying individuals based on them have been developed~\cite{fahmy2010online,galbally2015line,sanmorino2012survey,sesa2012information}.
% biometric
The use of biometric identity verification is growing with the popularity of mobile devices. Biometric identity verification systems are more secure, convinient, and provide greater accountability than traditional passwords~\cite{hutton2004biometrics,jain2006biometrics,ross2006handbook}.
% camera sensor
In-air signatures have become possible with the development of specialized sensors, such as depth cameras and magnetic sensors~\cite{bailador2011analysis, jeon2012system,ketabdar2012magnetic,malik20183dairsig}.
% WiFI CSI
Recently, the in-air signatures can also be acquired using the Wi-fi Channel State Information (``CSI'') signals. Typically the user makes a certain gesture when signing their signature on a piece of paper. With in-air signatures, the user simply makes the same gesture in the air. Their movements are detected by Wi-fi receivers with appropriate software installed. The software detects changes in the Wi-fi signals caused by the hand movements. This technique uses the existing widespread Wi-Fi devices, so it does not require the use of any special input devices. 
\cite{abdelnasser2015wigest,nandakumar2014wi} used CSI signals to recognize users' gestures and~\cite{moon2017air} used them to specifically identify signatures written in the air. 
However, due to limitations of the traditional feature extractor,~\cite{moon2017air} was only able to identify in-air signatures entering from a specific direction. This makes the user difficult to input their signatures.
% Deep learning
Recent studies have been conducted on the feasibility of using deep learning to characterize the CSI signals. Deep learning-based models are popular because of their automated feature extractors and the fact that they have superior classification capabilities than traditional models. Deep learning has been used to recognize users based on their vital sign~\cite{zhang2017pulseprint,szankin2018long} and behavior~\cite{shi2017smart,pokkunuru2018neuralwave}.
% our methods
In this thesis, the deep learning technology is used to create a Wi-Fi CSI system that can recognize a user's identity with multi-direction in-air signatures.
The deep learning model tested in this study is a triplet network where our task is to improve the model's convergence speed using the kernel and range space learning technique~\cite{toh2018gradient}.

%\newpage
\newpage
\section{Motivation and Contributions}
The main contributions of this thesis
\begin{itemize}
\item Proposing a system to verify user identities using Wi-Fi handwritten signature signals using a deep triplet network;
\item Using the kernel and range (KAR) space learning to mine distinctive triplet inputs which boost convergence speed and reduce triplet network training loss; and
\item Empirically testing the proposed system on a dataset of Wi-Fi handwritten signatures collected from 98 subjects.
\end{itemize}

\section{Paper Organization}
The rest of this thesis is organized as follows.  Section 2 discusses related works about triplet networks and KAR space learning. Section 3 discusses the proposed system. Section 4 describes this thesis's experimental and analysis results. Section 5 concludes the thesis.